---
title: 'HW_3'
author: "Louie Chen"
date: "September 30, 2018"
output: html_document
---

Tutorial to ggplots and other basic R knowledge: https://yijutseng.github.io/DataScienceRBook/vis.html#section-8.1

```{r}

#Import data

##Method 1
'library(gdata)                   # load gdata package 
help(read.xls)                  # documentation 
mydata = read.csv("D:/NTU_DataScience (R)/NTU_CSX_DataScience/Week_3/Course/metro-bike-share-trip-data.csv")
mydata'

##Method 2
library(readr)
metro_bike_share_trip_data <- read_csv("D:/NTU_DataScience (R)/NTU_CSX_DataScience/Week_3/Course/metro-bike-share-trip-data.csv")
'View(metro_bike_share_trip_data)'
library(ggplot2)
```

Duration distribution - bar chart
```{r}
ggplot(data = metro_bike_share_trip_data, aes(x = Duration)) +
geom_bar(fill = "lightblue", colour = "black")

```

Start Station Location Distribution - scatter chart
```{r}
'# remove NA in the sixth column in data
na.omit(metro_bike_share_trip_data[6])'
'# means first row and sixth column in data
metro_bike_share_trip_data[1,6]'

ggplot(data = metro_bike_share_trip_data, mapping = aes(x = Starting_Station_Longitude, y = Starting_Station_Latitude)) +
    geom_point(alpha = 0.005, color = "blue") + scale_y_continuous(limits = c(33.975, 34.07)) + scale_x_continuous(limits = c(-118.5, -118.2)) +   labs(title="Scatter Chart", 
           subtitle="Start Station Location Distribution", 
           caption="Source: Kaggle")

```

start station id distribution - Bar Chart
```{r}
theme_set(theme_bw())
ggplot(data = metro_bike_share_trip_data, aes(x = Starting_Station_ID)) +
geom_bar(width = 0.5, fill = "red", colour = "tomato2") + 
  scale_x_continuous(limits = c(3000, 3090)) + 
  labs(title="Bar Chart", 
           subtitle="Start Station Distribution", 
           caption="Source: Kaggle") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

Online source to only output certain range on plot: https://stackoverflow.com/questions/3606697/how-to-set-limits-for-axes-in-ggplot2-r-plots

start station id distribution - Bar Chart(Sorted)
```{r}
library(plyr)
count_star_sta_id <- count(metro_bike_share_trip_data, vars = "Starting_Station_ID")

count_star_sta_id <- count_star_sta_id[order(count_star_sta_id$freq), ]
count_star_sta_id <- na.omit(count_star_sta_id)

count_star_sta_id$Starting_Station_ID <- factor(count_star_sta_id$Starting_Station_ID, levels = count_star_sta_id$Starting_Station_ID)

ggplot(data = count_star_sta_id, aes(x = Starting_Station_ID, y = freq)) +
geom_bar(stat="identity", width = 0.5, fill = "red", colour = "tomato2") + 
  labs(title="Bar Chart (Sorted)", 
           subtitle="Start Station Distribution", 
           caption="Source: Kaggle") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

start station hot spot - Diverging Bars chart
```{r}
theme_set(theme_bw())  

# Data Prep
library(plyr)
count_star_sta_id <- count(metro_bike_share_trip_data, vars = "Starting_Station_ID")

count_star_sta_id <- na.omit(count_star_sta_id)

count_star_sta_id$freq <- round((count_star_sta_id$freq - mean(count_star_sta_id$freq))/sd(count_star_sta_id$freq), 2)  # compute normalized freq

count_star_sta_id <- count_star_sta_id[order(count_star_sta_id$freq), ]  # sort

type <- ifelse(count_star_sta_id$freq < 0, "below", "above")  # above / below avg flag

count_star_sta_id$Starting_Station_ID <- factor(count_star_sta_id$Starting_Station_ID, levels = count_star_sta_id$Starting_Station_ID)  # convert to factor to retain sorted order in plot.

# Diverging Barcharts
ggplot(count_star_sta_id, aes(x=Starting_Station_ID, y=freq, label=freq)) + 
  geom_bar(stat="identity", aes(fill = type), width=.5)  +
  scale_fill_manual(name="freq", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(subtitle="Start Station Hot Spot", 
       title= "Diverging Bars") + 
  coord_flip()
```

# Other things can be analyzed:
start time month growth - line chart
start time 點鐘 distribution - histogram
Passholder_Type duration histogram
starting station id, ending station id - slope chart
 
